---
title: "Practical One"
author: Fabio Iafrate
format: html
editor: visual
---

# Practical One

## Task 1

> Find all rows in “airquality” that have missing values.

This was completed by using the **complete.cases()** function. The **complete.cases()** function checks to see if the rows have missing (NA) values and returns *TRUE* if there are no missing values and *FALSE* otherwise.

```{r}
data(airquality) #load the data to see it better


###### 1. Rows With Missing Values ######

#complete cases checks if a row has any NA, returns false if NA
missing_rows <- airquality[!complete.cases(airquality), ] #Find all NA rows
missing_rows_numbers <- as.character(rownames(missing_rows)) #Stores as numbers 
print(missing_rows_numbers) # Prints to display 

```

The code outputs the row numbers with missing values. It is

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
